---
title: "Poisson_Point_Process_Generation"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is the initial step towards generating the Point Processes from Scratch. First, Homogenous Poisson Point Processes will be generated.

1. Generate a Homogenous Poisson Point Process on a Rectangle.
    A Poisson Point Process with an intensity of λ>0 on a bounded rectangular region, where the rectangle [0,w]×[0,h] with dimentions w>0 and h>0 is simulated. The area of the rectangle is A=hw. It is assumed that the bottom left corner of the rectangle is at the origin on the two dimentional space.

Number of Points on the rectangular region -  will follow a poisson random variable with the mean λA, as the number of points on the region not only depends on the intensity but also the size of the region. The function `rpois` is used to achieve this task.

Location of points - The points has to be positioned randomly using the cartesian coordinates. In a homogenous poisson point process, the x and y coordiantes of each point are independant and uniformly distributed. For the rectangle [0,w]×[0,h], the x coordinates are uniformly sampled from the interval [0,w] and y coordinates are uniformly sampled from interval [0,h]. If the bottom left corner of the rectangle is located at the (xo,yo), the random x and y coordinates needs to be shifted by adding xo and yo respectively to them. This is done using the `runif`.

```{r}
#Simulation area parameters
xMin <- 0;xMax <- 1
yMin <- 0;yMax <- 1
xDelta <- xMax-xMin; yDelta <- yMax-yMin #rectangle dimensions
areaTotal <- xDelta*yDelta
 
#Point process parameters
lambda <- 50 #intensity (ie mean density) of the Poisson process
 
#Simulate Poisson point process
numbPoints <- rpois(1,areaTotal*lambda)#Poisson number of points
xx <- xDelta*runif(numbPoints)+xMin#x coordinates of Poisson points
yy <- xDelta*runif(numbPoints)+yMin#y coordinates of Poisson points
 
#Plot
plot(xx,yy,'p',xlab='x',ylab='y',col='blue')
```
2. Generate a Homogenous Poisson Point Process on a Disk.
    A Poisson Point Process with an intensity of λ>0 on a bounded disk region with radius r>0 is simulated here. 
    
Number of Points - The number of points of a Poisson point process falling within a circle of radius r>0 is a Poisson random variable with mean  λA, where A=πr^2 is the area of the disk.This is done using the R function `rpois`. 

Locations of points - All the points need to be randomly and uniformly positioned on a disk. This can be done in polar coordinates.  The angular and radial coordinate respectively will be denoted by θ and ρ. To generate the random angular (or θ) values, an uniform random variables between zero and one will be produced.Then all these numbers will be multiplyed by 2π, meaning that all the numbers now fall between 0 and 2π.

To generate the random radial (or ρ) values, a reasonable guess would be to do the same as before and generate uniform random variables between zero and one, and then multiply them by the disk radius r. But that would be wrong.
  

```{r}
#Simulation area parameters
xMin <- 0;xMax <- 1
yMin <- 0;yMax <- 1
xDelta <- xMax-xMin; yDelta <- yMax-yMin #rectangle dimensions
areaTotal <- xDelta*yDelta
 
#Point process parameters
lambda <- 50 #intensity (ie mean density) of the Poisson process
 
#Simulate Poisson point process
numbPoints <- rpois(1,areaTotal*lambda)#Poisson number of points
xx <- xDelta*runif(numbPoints)+xMin#x coordinates of Poisson points
yy <- xDelta*runif(numbPoints)+yMin#y coordinates of Poisson points
 
#Plot
plot(xx,yy,'p',xlab='x',ylab='y',col='blue')
```
